Depth and thickness of tectonic tremor in the northeastern Olympic Peninsula

### 2021-06-22 paper/agujournaltemplate.tex
Correct paper

\begin{abstract}
Tectonic tremor has been explained as a swarm of low-frequency earthquakes, which are located on a narrow fault at the plate boundary. However, due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source and the distance of the source to the plate interface with great precision. The thickness of the tremor region is also not well constrained. The tremor may be located on a narrow fault like the low-frequency earthquakes appear to be, or distributed over a few kilometers wide low shear-wave velocity layer in the upper oceanic crust, which is thought to be a region with high pore-fluid pressure. Lag times of peaks in the cross-correlation of the horizontal and vertical components of tremor seismograms, recorded by small-aperture arrays in the Olympic Peninsula, Washington, are interpreted to be S minus P times.  Tremor depths are estimated from these S minus P times using epicenters from a tremor catalog. The tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick and is very close to the depths of low-frequency earthquakes. The tremor is distributed over a wider depth range than the low-frequency earthquakes. However, due to the uncertainty on the depth, it is difficult to conclude whether the source of the tremor is located at the top of the subducting oceanic crust, in the lower continental crust just above the plate boundary, or in a narrow zone at the plate boundary.
\end{abstract}

@@

Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor has been explained as a swarm of small, low frequency earthquakes (LFEs) ~\cite{SHE_2007_nature}, that is small magnitude earthquakes (M $\sim$ 1) where frequency content (1-10 Hz) is lower than for ordinary small earthquakes (up to 20 Hz). The source of the LFEs is located on the plate boundary, and their focal mechanisms represent shear slip on a low-angle thrust fault dipping in the same direction as the plate interface ~\cite{IDE_2007_GRL, ROY_2014}. Due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source with the same precision. The tremor source is assumed to be located close to the plate boundary. In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\cite{OBA_2002, ROG_2003}. Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS). \\

@@

The source of the fluids could be the dehydration of hydrous minerals within the subducting oceanic crust ~\cite{SHE_2006}. ~\citeA{FAG_2011} computed the equilibrium mineral assemblages at different P-T conditions, and compared it to the P-T path of the subducting oceanic crust in Shikoku and Cascadia. They noted that for most of the P-T path, there are no dehydration reactions and the slab remains fluid-absent, except for depths between 30 and 35 km depth for Shikoku, and depths between 30 and 40 km for Cascadia, where the mineral model predicts significant water release. These depth ranges coincide with the depth range where tremor has been observed. They concluded that abundant tremor activity requires metamorphic conditions where localized dehydration occurs during subduction, which explains why the generation of tectonic tremor is restricted to a small range of depth along the plate boundary. In subduction zones where dehydration reactions are more widely distributed, there would be a more diffuse pattern of tremor activity that would be harder to detect. \\

@@

In this study, we extend on the method used by ~\citeA{LAR_2009} using the cross-correlation between horizontal and vertical components of seismic recordings to estimate the depth of the source of the tectonic tremor, and the depth extent of the region from which tremor originates. If indeed tremor is made of swarms of LFEs, and both represent the regions of deformation during ETS events, we would expect the thickness of the tremor to be the same as the thickness of the LFEs. If not, tremor may be occurring where LFEs are harder to detect because either smaller in amplitude, or spread out over continuous space and not as clearly repeating, and thus harder to detect.

@@

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the one-minute long time windows corresponding to all the tremor epicenters located within the grid cell. For each three-component seismic station and each channel, we detrended the data, tapered the first and last 5 seconds with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross correlation functions over all the seismic stations of the array. We call this stack $S_{i j} (\tau)$ for time lag $\tau$, the $i$-th channel ($i$=1 or 2 for the North or East channel) and the $j$-th one-minute time window for a given epicenter grid/array pair. At each step of this method we experimented with a linear stack, a $n$th-root stack, and a phase-weighted stack ~\cite{SCH_1997} and found that the phase-weighted stack worked best. So, all the stacks discussed here used the phase-weighted method.  Figure 2 shows an example of the envelopes of $S_{i j} (\tau)$ for the Big Skidder array for the 110 one-minute long time windows when tremor was detected in the 5 km by 5 km grid cell located 7 kilometers southwest of the array. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we interpret this peak to correspond to the time lag between the arrival of the direct P- and S-waves.

@@

To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match the stacked cross-correlations well, and the ones that do not. In order to do the clustering, we stacked the stacked cross-correlation functions, $S_{i j} (\tau)$, over all the one-minute-long time windows to obtain $S_i (\tau)$. For each one-minute time window $j$ we cross-correlated $S_i (\tau)$ with $S_{i j} (\tau)$ to obtain $S^c_{i j} (\tau)$.  For each $i$ and $j$ we determine 3 numbers from $S^c_{i j} (\tau)$: its value at zero lag time, its maximum absolute value and the time lag at which it takes its maximum absolute value, and we determine one number from the $S_{i j} (\tau)$: the ratio of its maximum absolute value to its rms. We considered lag times from 12 to 14 seconds to compute the value of the rms. Each one-minute-long time window is thus associated with eight values of quality criteria, four for each of the the East and North channels. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose to have 2 clusters, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the envelopes of the cross-correlation functions over all the one-minute-long time windows belonging to that cluster using a phase-weighted stack. We tried using more than 2 clusters, but it did not improve the final stack of the best cluster. Figure 3 shows the stack of the envelopes of the $S_{i j} (\tau)$ over the time windows $j$ that correspond to cluster 0 (red) and cluster 1 (blue).  The clustering has greatly improved the amplitude of the peak for cluster 1, and made the peak nearly disappear for cluster 0.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/BS_-05_-05_PWS_PWS_cluster_stackcc.eps}
\caption{Stack of the cross-correlation functions over all the 110 time windows from Figure 2. The black line is the stack over all time windows of the stack of cross-correlation functions over all stations within the array. The red line on the left panels is the stack over all the time windows in the first cluster of the envelopes of the stack of cross-correlation functions over stations within the array.  This corresponds to time windows that do not fit well with the stack. The blue lines on the right panels are the same as the red lines, except it uses time windows in the second cluster (which contains the time windows that fit well with the stack). Top panels are the cross-correlation of the EW component with the vertical component, and bottom panels are the cross-correlation of the NS component with the vertical component. The stacked cross-correlation function has been cut between 2 and 8 seconds to focus on the time lags around the peak.}
\label{pngfiguresample}
\end{figure}

@@

Finally, the data from some of the arrays are very noisy, which makes it hard to see a signal emerging when stacking over the one-minute-long time windows. We chose to keep only the locations for which the ratio between the maximum value of the stack of the envelopes of the cross-correlation functions to the root mean square is higher than 5. We compute the root mean square of the cross-correlation for a time lag between 12 and 14s because we do not expect any reflected wave to arrive that late after a direct wave. The corresponding stack of the envelopes of the cross-correlation functions are shown in Figure 4 for the East-West component and the North-South component. The grey vertical line shows the theoretical time lag between the arrival of the direct P-wave and the arrival of the direct S-wave using the corresponding 1D velocity model for this array and the plate boundary model from ~\citeA{MCC_2006}, the dashed line corresponds to the slab model from ~\citeA{PRE_2003}. There is a good agreement between the timing of the peak and the theoretical time lag for most of the locations of the source of the tremor.

@@

To estimate the uncertainty on the depth of the source of the tremor, we computed the width of the stack of the envelopes of the cross-correlation functions at half the maximum of the peak. We then computed the associated depth difference using the 1D velocity model. As we computed the envelopes of the cross-correlation functions before stacking, the peak is large and also the associated uncertainty on the depth. Nevertheless, about 75\% of the data points have an uncertainty under 8km (Figure 5). The uncertainties are higher for the Burnt Hill array, and are almost always lower than 8km for the five other arrays. We tried different stacking methods, both for the stacking of the correlation functions over all the seismic stations of a given array, and for the stacking of the envelopes of those stacks over the one-minute-long time windows when tremor is detected. We experimented with a linear stack, a nth-root stack, and a phase-weighted stack, and found that using a phase-weighted stack for both the stacking over the stations and the stacking over one-minute-long time windows gave the lower uncertainty on the depth of the source of the tremor.

@@

In the following, we kept only the data points for which the maximum amplitude of the stack is higher than 0.05. Figures 6 and 7 show respectively a map and several cross-sections of the depth of the source of the tremor. Figure 8 shows a map of the distance between the source of the tremor and the plate boundary from the Preston model, alongside with the depth of the low-frequency earthquake families observed by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. There is a good agreement between the depth of the source of the tremor, the depth of the low-frequency earthquake families, and the depth of the plate boundary. On the cross-section figures, we compare the distance of the source of the tremor to the plate boundary for different locations of the cross-section. We note that the depth of the tremor matches well the depth of the low-frequency earthquakes.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/depth_PWS_PWS.eps}
\caption{Map of the depth of the source of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The three black lines indicate the positions of the cross-sections shown in Figures 7 to 9. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles)}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=5cm, trim={4.5cm 0.5cm 6cm 0cm},clip, angle=270]{figures/section_strike_PWS_PWS.eps}
\caption{Cross-sections showing depth of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR} for cross sections A-A' , B-B', and C-C' shown in Figure 6. The black line corresponds to the plate boundary profile of McCrory (2006) along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of Preston (2003). The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

@@

Previous studies have shown evidence for seismic anisotropy near the slab interface ~\cite{NIK_2009} and in the overriding continental crust ~\cite{CAS_1996}. If the source of the tremor is located in or under the anisotropic layer, the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave could be different different depending on whether we consider the East-West component or the North-South component. To verify whether possible anisotropy could affect our results, we computed the time difference between the timing of the maximum of the stack of the envelopes of the stacked cross-correlation functions between the East-West component and the vertical component, and between the North-South component and the vertical component. We also computed the associated difference in depth of the source of the tremor (see Figure 12) for all arrays and all locations of the tremor. The difference in time is generally less than 0.25 s while the difference in depth is generally less than 2 kilometers, which is similar to the uncertainty on the tremor depth. As the time difference between the East-West component and the North-South component due to anisotropy is expected to vary depending on the relative position of the array compared to the source of the tremor, we also plotted the time difference as a function of the distance from source to array, and as a function of the azimuth, as well as the corresponding depth difference (Figure 12). The time difference does not increase with the distance from the source to the array, and there is no seismic path orientation that gives bigger time difference. Thus, seismic anisotropy does not seem to have a significant effect on the time lags between the arrival of the direct P-wave and the arrival of the direct S-wave, and can be neglected. ~\citeA{NIK_2009} observed anisotropy in the low-velocity layer beneath station GNW, located in the eastern Olympic Peninsula, south of the eight seismic arrays used in this study. If the tremor is located above the low-velocity layer, we do not expect this source of seismic anisotropy to introduce a significant effect on our measured time lags. ~\citeA{NIK_2009} locate the low-velocity layer above the plate boundary, in the lower continental crust, which is shallower than the location given by ~\citeA{BOS_2013}, who locates the low-velocity layer in the upper oceanic crust. However, even if a low-velocity layer with 5 \% anisotropy is located in the lower continental crust as indicated by ~\citeA{NIK_2009}, the resulting difference in time lags should not be more than a few tenths of a second, and the resulting difference in depth should not be more than 2 kilometers. ~\citeA{CAS_1996} observed anisotropy in the continental crust, especially in the upper 20 kilometers to the north of our arrays. However, the anisotropy resulted in time delays of the seismic waves of no more than 0.32s for deep earthquakes (40-60 km depth) and not more than 0.20s for shallow earthquakes (15-30 km depth). The corresponding difference in tremor depth is similar to the uncertainty.

@@

Several approximations have been made to compute the tremor depth. First, we stacked over all the tremor in 5km by 5km grid cell. The depth difference corresponding to epicentral tremor locations at the center versus the edge of a cell varies from about 1 to 2 km for cells near the array to cells 25 km from an array. Second, the VP / VS ratio is not very well known. A variation of about 1\% of the VP / VS ratio used to compute the depth  would lead to a variation of about 1-1.5 km of the corresponding tremor depth (Figure 13).

@@

As proposed by ~\citeA{KAO_2009}, the source of the tremor could be distributed inside a layer that is 15 or more kilometers thick. We have assumed that all the tremor within a given grid cell originate from the same depth, and averaged over all the data to get the tremor depth. However, instead of being located on the same plane near the plate boundary, the tremor may be scattered over a layer surrounding the plate boundary. To compute the thickness of this layer, for each location of the array and the source of the tremor, we computed for each one-minute-long time window for which the cross-correlation function matches well the stacked cross-correlation the time lag between the time corresponding to the maximum absolute value for the cross-correlation function and the time corresponding to the maximum absolute value for the stacked cross-correlation. We thus obtained a distribution of time lags and the corresponding distribution of depths, and we tried to estimate the scale of the interval over which the depths vary. Using the standard deviation could lead to overestimate the width of the interval, and the corresponding thickness of the tremor layer, as the standard deviation is very sensitive to outliers. Instead, we use the Qn estimator of ~\citeA{ROU_1993}, which is a more robust estimator of scale, similar to the median absolute deviation. We then used the scale of the interval over which the tremor depth varies as an estimate of the thickness of the tremor layer (see Figure 14). The average thickness of the tremor zone is about 1.6 km, while nearly all the values of the thickness are less than 3km, which is substantially lower than the depth extent from ~\citeA{KAO_2009}. The smallest values for the thickness are about 0.8-2 kilometers.

@@

Another way of estimating the thickness of the tremor zone is to fit a plane with a linear regression from all the values of the depth and compute the residuals from the regression (Figure 15). The Qn estimator of scale ~\cite{ROU_1993} for the error between the tremor depth and the fitted plane is only 1.3 km. \\

The median of the absolute value of the difference between the depth of the plate boundary and the depth of the tremor is 1.92 km for the McCrory model and 1.88 km for the Preston model. The median of the difference between the depth of the plate boundary and the depth of the tremor is -0.85 km for the McCrory model, and -0.21 km for the Preston model, which locates the tremor at the plate boundary or a little bit above. The Qn estimator of scale for these models is 1.39 and 1.31 km. So, to a very good approximation,  the tremor appears to coincide with the plate boundary. \\

By comparison, Qn estimator of scale for the low-frequency earthquakes is 0.99 km for the McCrory model and 0.87 km for the Preston model. The median of the absolute value of the difference between the depth of the plate boundary and the depth of the LFEs is 1.22 km for the McCrory model and 0.96 km for the Preston model. The median of the difference between the depth of the plate boundary and the depth of the LFEs is -0.49 km for the McCrory model and -0.19 km for the Preston model, which locates the LFEs at the plate boundary or a little bit above.

\begin{table}
\caption{Summary of distances to plate boundary}
\centering
\begin{tabular}{l c c c}
\hline
 & Median & MAD & Q \\
\hline
Tremor depth - Preston depth & - 0.207 km & 2.623 km & 1.313 km \\
Tremor depth - McCrory depth & - 0.850 km & 2.429 km & 1.390 km \\
LFE depth - Preston depth & - 0.189 km & 1.312 km & 0.871 km \\
LFE depth - McCrory depth & - 0.495 km & 1.805 km & 0.988 km \\
\hline
\end{tabular}
\end{table}

@@

>>>>

\begin{abstract}
Tectonic tremor has been explained as a swarm of low-frequency earthquakes, which are located on a narrow fault at the plate boundary. However, due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source and the distance of the source to the plate interface with great precision. The thickness of the tremor region is also not well constrained. The tremor may be located on a narrow fault as the low-frequency earthquakes appear to be, or distributed over a few kilometers wide low shear-wave velocity layer in the upper oceanic crust, which is thought to be a region with high pore-fluid pressure. Lag times of peaks in the cross-correlation of the horizontal and vertical components of tremor seismograms, recorded by small-aperture arrays in the Olympic Peninsula, Washington, are interpreted to be S minus P times.  Tremor depths are estimated from these S minus P times using epicenters from a tremor catalog. The tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick and is very close to the depths of low-frequency earthquakes. The tremor is distributed over a wider depth range than the low-frequency earthquakes. However, due to the uncertainty on the depth, it is difficult to conclude whether the source of the tremor is located at the top of the subducting oceanic crust, in the lower continental crust just above the plate boundary, or in a narrow zone at the plate boundary.
\end{abstract}

@@

Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor has been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\cite{SHE_2007_nature}, that is small magnitude earthquakes (M $\sim$ 1) where frequency content (1-10 Hz) is lower than for ordinary small earthquakes (up to 20 Hz). The source of the LFEs is located on the plate boundary, and their focal mechanisms represent shear slip on a low-angle thrust fault dipping in the same direction as the plate interface ~\cite{IDE_2007_GRL, ROY_2014}. Due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source with the same precision. The tremor source is assumed to be located close to the plate boundary. In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\cite{OBA_2002, ROG_2003}. Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS). \\

@@

The source of the fluids could be the dehydration of hydrous minerals within the subducting oceanic crust ~\cite{SHE_2006}. ~\citeA{FAG_2011} computed the equilibrium mineral assemblages at different P-T conditions, and compared it to the P-T path of the subducting oceanic crust in Shikoku and Cascadia. They noted that for most of the P-T path, there are no dehydration reactions and the slab remains fluid-absent, except for depths between 30 and 35 km for Shikoku, and depths between 30 and 40 km for Cascadia, where the mineral model predicts significant water release. These depth ranges coincide with the depth range where tremor has been observed. They concluded that abundant tremor activity requires metamorphic conditions where localized dehydration occurs during subduction, which explains why the generation of tectonic tremor is restricted to a small range of depth along the plate boundary. In subduction zones where dehydration reactions are more widely distributed, there would be a more diffuse pattern of tremor activity that would be harder to detect. \\

@@

In this study, we extend on the method used by ~\citeA{LAR_2009} using the cross-correlation between horizontal and vertical components of seismic recordings to estimate the depth of the source of the tectonic tremor, and the depth extent of the region from which tremor originates. If indeed tremor is made of swarms of LFEs, and both represent the regions of deformation during ETS events, we would expect the thickness of the tremor to be the same as the thickness of the LFEs. If not, tremor may be occurring where LFEs are harder to detect because either smaller in amplitude, or spread out over continuous space and not as clearly repeating.

@@

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the one-minute-long time windows corresponding to all the tremor epicenters located within the grid cell. For each three-component seismic station and each channel, we detrended the data, tapered the first and last 5 seconds with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross correlation functions over all the seismic stations of the array. We call this stack $S_{i j} (\tau)$ for time lag $\tau$, the $i$-th channel ($i$=1 or 2 for the North or East channel) and the $j$-th one-minute-long time window for a given epicenter grid/array pair. At each step of this method we experimented with a linear stack, a $n$th-root stack, and a phase-weighted stack ~\cite{SCH_1997} and found that the phase-weighted stack worked best. So, all the stacks discussed here used the phase-weighted method.  Figure 2 shows an example of the envelopes of $S_{i j} (\tau)$ for the Big Skidder array for the 110 one-minute-long time windows when tremor was detected in the 5 km by 5 km grid cell located 7 kilometers southwest of the array. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we interpret this peak to correspond to the time lag between the arrival of the direct P- and S-waves. \\

@@

To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match the stacked cross-correlations well, and the ones that do not. In order to do the clustering, we stacked the stacked cross-correlation functions, $S_{i j} (\tau)$, over all the one-minute-long time windows to obtain $S_i (\tau)$. For each one-minute time window $j$ we cross-correlated $S_i (\tau)$ with $S_{i j} (\tau)$ to obtain $S^c_{i j} (\tau)$. For this cross-correlation, we keep only the values of lag times $\tau$ between one second before and one second after the time of the maximum of the absolute value of the stack $S_i (\tau)$. For each $i$ and $j$ we determine 3 numbers from $S^c_{i j} (\tau)$: its value at zero lag time, its maximum absolute value and the time lag at which it takes its maximum absolute value, and we determine one number from $S_{i j} (\tau)$: the ratio of its maximum absolute value to its rms. We considered lag times $\tau$ from 12 to 14 seconds to compute the value of the rms of the $S_{i j} (\tau)$. Each one-minute-long time window is thus associated with eight values of quality criteria, four for each of the the East and North channels. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose to have 2 clusters, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the envelopes of the cross-correlation functions over all the one-minute-long time windows belonging to that cluster using a phase-weighted stack. We tried using more than 2 clusters, but it did not improve the final stack of the best cluster. Figure 3 shows the stack of the envelopes of the $S_{i j} (\tau)$ over the time windows $j$ that correspond to cluster 0 (red) and cluster 1 (blue).  The clustering has greatly improved the amplitude of the peak for cluster 1, and made the peak nearly disappear for cluster 0. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/BS_-05_-05_PWS_PWS_cluster_stackcc.eps}
\caption{Stack of the cross-correlation functions over all the 110 time windows from Figure 2. The black line is the stack $S_i (\tau)$ over all time windows of the stack of cross-correlation functions over all stations within the array. The red line on the left panels is the stack over all the time windows in cluster 0 of the envelopes of the stack of cross-correlation functions over stations within the array.  This corresponds to time windows that do not fit well with the stack. The blue lines on the right panels are the same as the red lines, except it uses time windows in cluster 1 (which contains the time windows that fit well with the stack). Top panels are the cross-correlation of the EW component with the vertical component, and bottom panels are the cross-correlation of the NS component with the vertical component. The stacked cross-correlation function has been cut between 2 and 8 seconds to focus on the time lags around the peak.}
\label{pngfiguresample}
\end{figure}

@@

Finally, the data from some of the arrays are very noisy, which makes it hard to see a signal emerging when stacking over the one-minute-long time windows. We chose to keep only the locations for which the ratio between the maximum value of the stack of the envelopes of the cross-correlation functions to the root mean square is higher than 5. We compute the root mean square of the cross-correlation for a time lag between 12 and 14s because we do not expect any reflected wave to arrive that late after a direct wave. The corresponding stack of the envelopes of the cross-correlation functions are shown in Figure 4 for the East-West component and the North-South component. The grey vertical line shows the theoretical time lag between the arrival of the direct P-wave and the arrival of the direct S-wave using the corresponding 1D velocity model for this array and the plate boundary model from ~\citeA{MCC_2006}, the dashed line corresponds to the slab model from ~\citeA{PRE_2003}. There is a good agreement between the timing of the peak and the theoretical time lag for most of the locations of the source of the tremor. \\

@@

To estimate the uncertainty on the depth of the source of the tremor, we computed the width of the stack of the envelopes of the cross-correlation functions at half the maximum of the peak. We then computed the associated depth difference using the 1D velocity model. As we computed the envelopes of the cross-correlation functions before stacking, the peak is large and also the associated uncertainty on the depth. Nevertheless, about 75\% of the data points have an uncertainty under 8km (Figure 5). The uncertainties are higher for the Burnt Hill array, and are almost always lower than 8km for the five other arrays. We tried different stacking methods, both for the stacking of the correlation functions over all the seismic stations of a given array, and for the stacking of the envelopes of those stacks over the one-minute-long time windows when tremor is detected. We experimented with a linear stack, a $n$th-root stack, and a phase-weighted stack, and found that using a phase-weighted stack for both the stacking over the stations and the stacking over one-minute-long time windows gave the lower uncertainty on the depth of the source of the tremor. \\

@@

In the following, we kept only the data points for which the maximum amplitude of the stack is higher than 0.05. Figures 6 and 7 show respectively a map and several cross-sections of the depth of the source of the tremor. Figure 8 shows a map of the distance between the source of the tremor and the plate boundary from the Preston model, alongside with the depth of the low-frequency earthquake families observed by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. There is a good agreement between the depth of the source of the tremor, the depth of the low-frequency earthquake families, and the depth of the plate boundary. On the cross-section figures, we compare the distance of the source of the tremor to the plate boundary for different locations of the cross-section. We note that the depth of the tremor matches well the depth of the low-frequency earthquakes. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/depth_PWS_PWS.eps}
\caption{Map of the depth of the source of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The three black lines indicate the positions of the cross-sections shown in Figure 7. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles)}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=5cm, trim={4.5cm 0.5cm 6cm 0cm},clip, angle=270]{figures/section_strike_PWS_PWS.eps}
\caption{Cross-sections showing depth of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR} for cross sections A-A' , B-B', and C-C' shown in Figure 6. The black line corresponds to the plate boundary profile of McCrory (2006) along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of Preston (2003). The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less than 10 km away from the profile line are shown. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

@@

Previous studies have shown evidence for seismic anisotropy near the slab interface ~\cite{NIK_2009} and in the overriding continental crust ~\cite{CAS_1996}. If the source of the tremor is located in or under the anisotropic layer, the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave could be different different depending on whether we consider the East-West component or the North-South component. To verify whether possible anisotropy could affect our results, we computed the time difference between the timing of the maximum of the stack of the envelopes of the stacked cross-correlation functions between the East-West component and the vertical component, and between the North-South component and the vertical component. We also computed the associated difference in depth of the source of the tremor (see Figure 9) for all arrays and all locations of the tremor. The difference in time is generally less than 0.25 s while the difference in depth is generally less than 2 kilometers, which is similar to the uncertainty on the tremor depth. As the time difference between the East-West component and the North-South component due to anisotropy is expected to vary depending on the relative position of the array compared to the source of the tremor, we also plotted the time difference as a function of the distance from source to array, and as a function of the azimuth, as well as the corresponding depth difference (Figure 9). The time difference does not increase with the distance from the source to the array, and there is no seismic path orientation that gives bigger time difference. Thus, seismic anisotropy does not seem to have a significant effect on the time lags between the arrival of the direct P-wave and the arrival of the direct S-wave, and can be neglected. ~\citeA{NIK_2009} observed anisotropy in the low-velocity layer beneath station GNW, located in the eastern Olympic Peninsula, south of the eight seismic arrays used in this study. If the tremor is located above the low-velocity layer, we do not expect this source of seismic anisotropy to introduce a significant effect on our measured time lags. ~\citeA{NIK_2009} locate the low-velocity layer above the plate boundary, in the lower continental crust, which is shallower than the location given by ~\citeA{BOS_2013}, who locates the low-velocity layer in the upper oceanic crust. However, even if a low-velocity layer with 5 \% anisotropy is located in the lower continental crust as indicated by ~\citeA{NIK_2009}, the resulting difference in time lags should not be more than a few tenths of a second, and the resulting difference in depth should not be more than 2 kilometers. ~\citeA{CAS_1996} observed anisotropy in the continental crust, especially in the upper 20 kilometers to the north of our arrays. However, the anisotropy resulted in time delays of the seismic waves of no more than 0.32s for deep earthquakes (40-60 km depth) and not more than 0.20s for shallow earthquakes (15-30 km depth). The corresponding difference in tremor depth is similar to the uncertainty. \\

@@

Several approximations have been made to compute the tremor depth. First, we stacked over all the tremor in 5km by 5km grid cell. The depth difference corresponding to epicentral tremor locations at the center versus the edge of a cell varies from about 1 to 2 km for cells near the array to cells 25 km from an array. Second, the Vp / Vs ratio is not very well known. A variation of about 1\% of the Vp / Vs ratio used to compute the depth  would lead to a variation of about 1-1.5 km of the corresponding tremor depth (Figure 10). \\

@@

As proposed by ~\citeA{KAO_2009}, the source of the tremor could be distributed inside a layer that is 15 or more kilometers thick. We have assumed that all the tremor within a given grid cell originate from the same depth, and averaged over all the data to get the tremor depth. However, instead of being located on the same plane near the plate boundary, the tremor may be scattered over a layer surrounding the plate boundary. To compute the thickness of this layer, for each location of the array and the source of the tremor, we computed for each one-minute-long time window for which the cross-correlation function matches well the stacked cross-correlation the time lag between the time corresponding to the maximum absolute value for the cross-correlation function and the time corresponding to the maximum absolute value for the stacked cross-correlation. We thus obtained a distribution of time lags and the corresponding distribution of depths, and we tried to estimate the scale of the interval over which the depths vary. Using the standard deviation could lead to overestimate the width of the interval, and the corresponding thickness of the tremor layer, as the standard deviation is very sensitive to outliers. Instead, we use the Qn estimator of ~\citeA{ROU_1993}, which is a more robust estimator of scale, similar to the median absolute deviation. We then used the scale of the interval over which the tremor depth varies as an estimate of the thickness of the tremor layer (see Figure 11). The average thickness of the tremor zone is about 1.6 km, while nearly all the values of the thickness are less than 3km, which is substantially lower than the depth extent from ~\citeA{KAO_2009}. The smallest values for the thickness are about 0.8-2 kilometers. \\

@@

Another way of estimating the thickness of the tremor zone is to fit a plane with a linear regression from all the values of the depth and compute the residuals from the regression (Figure 12). The Qn estimator of scale ~\cite{ROU_1993} for the error between the tremor depth and the fitted plane is only 1.3 km. \\

@@

In Table 1, we report the values of the median, the median absolute deviation (MAD), and the Qn estimator of scale for the difference between the depth of the tremor and the low-frequency earthquakes and the depth of the plate boundary, for the McCrory model and for the Preston model. Small negative values of the median correspond to tremor and low-frequency earthquakes located a little bit above the plate boundary. The depth range of the tremor is larger than the depth range of the low-frequency earthquakes, but it may be due to the uncertainty on the determination of the tremor depth. The tremor and low-frequency earthquakes look closer to the plate boundary for the Preston model than for the McCrory model. This may be due because the P-wave velocity model of ~\citeA{PRE_2003} was used as the initial velocity model by ~\citeA{MER_2020} to compute the full P- and S-wave velocity model that we use to get tremor depth from S minus P times. The Preston P-wave velocity model was also used by ~\citeA{CHE_2017_JGR} to locate the low-frequency earthquake families. \\

\begin{table}
\caption{Summary of distances to plate boundary}
\centering
\begin{tabular}{l c c c}
\hline
 & Median & MAD & Qn \\
\hline
Tremor depth - Preston depth & - 0.207 km & 2.623 km & 1.313 km \\
Tremor depth - McCrory depth & - 0.850 km & 2.429 km & 1.390 km \\
LFE depth - Preston depth & - 0.189 km & 1.312 km & 0.871 km \\
LFE depth - McCrory depth & - 0.495 km & 1.805 km & 0.988 km \\
\hline
\end{tabular}
\end{table}
